### 2008-11-20 docs/radial_base/radial_base.tex
more details, added Mike's fig

As there is no alignment between the 3-d Cartesian grid and the 1-d
spherical base state grid, we are free to choose their grid spacings
independently.  Experimentation has shown that the best choice results
from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
than the 1-d spherical base state spacing, $\Delta r$.  A typical simulation
may use $5 \Delta r = \Delta x$.
We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
complementary procedure that takes us from 3-d to 1-d {\tt average}.
\section{\tt fill\_3d}
Figure XXX shows the Cartesian grid with the spherical
base state overlayed (for simplicity, the figure is drawn in 2-d).  
The {\tt fill\_3d} procedure computes the radius of the cell center
from the center of the star,
\begin{equation}
r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
\end{equation}
where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
/ \Delta r)$ (here, our convention is to use 0-based indexing for the
base state).  We can then initialize a Cartesian cell quantity $q$ from its
corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
\subsection{\tt put\_1d\_array\_on\_cart}
\subsection{\tt put\_w0\_on\_edges}
\section{\tt average}
For the {\tt average} process, we again compute the radius of the
Cartesian cell center, and compute the index, $n$, of the
corresponding base state bin as above.  $q_{0,n}$ is then the average
of all the $q_{i,j,k}$ whose Cartesian cell center maps into radial
bin $n$.  However, if $\Delta r < \Delta x$, there may be some base
state bins that are not filled.  In this case, we will fill them via
interpolation. \MarginPar{more detail will be needed here.}

>>>>

As there is no alignment between the 3-d Cartesian grid and the 1-d
spherical base state grid, we are free to choose their grid spacings
independently.  Experimentation has shown that the best choice results
from picking the 3-d Cartesian grid spacing, $\Delta x$ to be coarser
than the 1-d spherical base state spacing, $\Delta r$.
A typical simulation may use $5 \Delta r = \Delta x$.
We call the procedure to go from 1-d to 3-d {\tt fill\_3d} and the
complementary procedure that takes us from 3-d to 1-d {\tt average}.
\section{\tt fill\_3d}
There are three different {\tt fill\_3d} modes:
\begin{enumerate}
\item The 1D array is cell-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian cell-centers.
\item The 1D array is edge-centered, and we map to Cartesian edges.  
\end{enumerate}
\subsection{1D Cell-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
Figure \ref{fig:mapping} shows the Cartesian grid with the spherical
base state overlayed (for simplicity, the figure is drawn in 2-d).  
We compute the radius of the cell center from the center of the star,
\begin{equation}
r = \sqrt{(x_i - x_c)^2 + (y_j - y_c)^2 + (z_k - z_c)^2}
\end{equation}
where $(x_c, y_c, z_c)$ are the coordinates of the center of the star.
We use this radius to find the corresponding radial bin as $n = \mathtt{int}(r
/ \Delta r)$ (here, our convention is to use 0-based indexing for the
base state).  We can then initialize a Cartesian cell quantity $q$ from its
corresponding base state quantity, $q_0$ as $q_{i,j,k} = q_{0,n}$.
\subsection{1D Edge-Centered to Cartesian Cell-Centered}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
\subsection{1D Edge-Centered to Cartesian Edges}
This mapping is done in the subroutine {\tt put\_1d\_array\_on\_cart}.
This case is only used to create {\tt w0mac} from {\tt w0} in the 
subroutine, {\tt put\_w0\_on\_edges}.
\section{\tt average}
For the {\tt average} process, we first create a coarse 1D radial array 
with $\Delta x = \Delta r$.  Then, we compute the radius of the Cartesian cell 
center, and compute the index, $n$, of the corresponding coarse radial 
array bin.  $q_{0,n}$ is then the average of all the $q_{i,j,k}$ whose 
Cartesian cell center maps into coarse radial bin $n$.  Then, we 
interpolate the coarse radial array onto the finer radial array.

\clearpage

\begin{figure}[tpb]
\begin{center}
\epsffile{./radial_base/mapping.eps}
\caption{\label{fig:mapping} The Cartesian grid and spherical base
state (shown here in 2-d for simplicity).  Here we represent the
spherical base state as concentric shells (blue lines).  Since the
base state is not aligned with the Cartesian grid we use on which
we discretize the star, we need to map between the two configurations.
The `$+$' symbols represent the Cartesian zone centers.  In our
mapping, the zone marked with the red symbol maps into the gray-shaded
radial bin. }
\label{fig:mapping}
\end{center}
\end{figure}
